---
title: "Daily R"
format: pdf
editor: visual
---

## Third Week (Sept. 23 \~ Sept. 29)

-   25th practice

    ```{r}
    library(ggplot2)
    library(dplyr)
    library(readxl)
    library(haven)

    kow <- read_sav("C:/Users/user/OneDrive - University of Illinois - Urbana/R/R/Doit_R-master/Doit_R-master/Data/Koweps_hpc10_2015_beta1.sav")

    list_job <- read_excel("C:/Users/user/OneDrive - University of Illinois - Urbana/R/R/Doit_R-master/Doit_R-master/Data/Koweps_codebook.xlsx", col_names = T, sheet = 2)

    kow <- rename(kow,
                  sex = h10_g3,
                  birth = h10_g4,
                  marriage = h10_g10,
                  religion = h10_g11,
                  income = p1002_8aq1,
                  code_job = h10_eco9,
                  code_region = h10_reg7)

    head(list_job)
    dim(list_job)

    kow <- left_join(kow, list_job, by = "code_job")

    kow %>% 
      filter(!is.na(code_job)) %>% 
      select(code_job, job) %>% 
      head(10)


    job_income <- kow %>% 
      filter(!is.na(job) &!is.na(income)) %>% 
      group_by(job) %>% 
      summarise(mean_income = mean(income)) ; job_income

    top10 <- job_income %>% 
      arrange(desc(mean_income)) %>% 
      head(10)
    top10

    ggplot(data = top10, aes(x = reorder(job, mean_income), y = mean_income)) +
      geom_col() +
      coord_flip()

    bottom10 <- job_income %>% 
      arrange(mean_income) %>% 
      head(10) 
    bottom10

    ggplot(data = bottom10, aes(x = reorder(job, -mean_income),
                                y = mean_income)) +
      geom_col() +
      coord_flip() +
      ylim(0, 850)


    ```

-   24th practice

    ```{r}

    library(ggplot2)
    library(dplyr)

    kow <- kow %>% 
      mutate(ageg = ifelse(age < 30, "young",
                           ifelse(age <=59, "middle", "old")))

    table(kow$ageg)

    ageg_income <- kow %>% 
      filter(!is.na(income)) %>% 
      group_by(ageg) %>% 
      summarise(mean_income = mean(income))

    ageg_income

    ggplot(data = ageg_income, aes(x = ageg, y = mean_income)) +
      geom_col()

    #in alphabetical order

    ggplot(ageg_income, aes(x = ageg, y = mean_income)) +
      geom_col() +
      scale_x_discrete(limits = c("young", "middle", "old"))


    #age group, gender, income

    names(kow)
    sex_income <- kow %>% 
      filter(!is.na(income)) %>% 
      group_by(ageg, sex) %>% 
      summarise(mean_income = mean(income))

    sex_income

    ggplot(data = sex_income, aes(x = ageg, y = mean_income, fill = sex)) +
      geom_col() +
      scale_x_discrete(limits = c("young", "middle", "old"))

    ggplot(data = sex_income, aes(x = ageg, y = mean_income, fill = sex)) +
      geom_col(position = "dodge") +
      scale_x_discrete(limits = c("young", "middle", "old"))


    sex_age <- kow %>% 
      filter(!is.na(income)) %>% 
      group_by(age, sex) %>% 
      summarise(mean_income = mean(income))

    ggplot(data = sex_age, aes(x = age, y = mean_income, col = sex)) +
      geom_line()
    ```

-   23rd making graph (Line Chart, box plot)

    ```{r}
    library(ggplot2)
    library(dplyr)


    data(economics)

    ggplot(data = economics, aes(x = date, y = unemploy)) + 
      geom_line()

    ggplot(data = economics, aes(x = date, y = psavert)) + 
      geom_line()

    ggplot(data = mpg, aes(x = drv, y = hwy)) + 
      geom_boxplot()

    mpg3 <- mpg %>% 
      filter(class %in% c("compact", "subcompact", "suv"))

    ggplot(data = mpg3, aes(x = class, y = cty)) +
      geom_boxplot()

    names(Koweps_hpc10_2015_beta1)

    library(foreign)
    library(dplyr)
    library(ggplot2)
    library(readxl)

    kow <- Koweps_hpc10_2015_beta1
    head(kow)
    str(kow)

    kow <- rename(kow,
                  sex = h10_g3,
                  birth = h10_g4,
                  marriage = h10_g10,
                  religion = h10_g11,
                  income = p1002_8aq1,
                  code_job = h10_eco9,
                  code_region = h10_reg7)

    class(kow$sex)
    table(kow$sex)

    kow$sex <- ifelse(kow$sex == 1, "male", "female")
    table(kow$sex)

    class(kow$income)
    summary(kow$income)
    qplot(kow$income)
    qplot(kow$income) + xlim(0, 1000)

    kow$income <- ifelse(kow$income %in% c(0, 9999), NA, kow$income)
    table(is.na(kow$income))

    sex_income <- kow %>% 
      filter(!is.na(income)) %>% 
      group_by(sex) %>% 
      summarise(mean_income = mean(income))
    sex_income

    ggplot(data = kow, aes(x = sex, y = income)) + 
      geom_col()

    #age and income

    class(kow$birth)
    qplot(kow$birth)

    summary(kow$birth)
    kow$age <- 2015 - kow$birth + 1
    summary(kow$age)

    age_income <- kow %>% 
      filter(!is.na(income)) %>% 
      group_by(age) %>% 
      summarise(mean_income = mean(income))

    head(age_income)
    ggplot(data = age_income, aes(x = age, y= mean_income)) +
      geom_line()
    ```

## Second Week (Sept. 16 \~ Sept. 22)

-   22nd making graphs (scatter plot, bar chart)

    ```{r}
    library(ggplot2)
    library(dplyr)
    data(mpg)

    ggplot(data = mpg, aes(x = displ, y = hwy)) + #data, aesthetic mappings
      geom_point() + #kinds of graph
      xlim(3, 6) +
      ylim(10, 30)

    ggplot(data = mpg, aes(x = cty, y = hwy)) + 
      geom_point()

    ggplot(data = midwest, aes(x = poptotal, y = popasian)) +
      geom_point() + 
      xlim(0, 500000) +
      ylim(0, 10000)

    #bar chart
    df_mpg <- mpg %>% 
      group_by(drv) %>% 
      summarise(mean_hwy = mean(hwy))
    df_mpg

    ggplot(data = df_mpg, aes(x = drv, y = mean_hwy)) +
      geom_col()

    ggplot(data = df_mpg, aes(x = reorder(drv, -mean_hwy), y = mean_hwy)) + 
             geom_col()

    ggplot(data = mpg, aes(x = drv)) + 
      geom_bar()

    ggplot(data = mpg, aes(x = hwy)) + 
      geom_bar()

    names(mpg)

    suv_mpg <- mpg %>% 
      group_by(manufacturer) %>% 
      filter(class == "suv") %>% 
      summarise(mean_cty = mean(cty))

    ggplot(data = suv_mpg, aes(x = manufacturer, y = mean_cty)) + 
      geom_col()

    ggplot(data = suv_mpg, aes(x = reorder(manufacturer, -mean_cty), y = mean_cty)) + geom_col()


    ggplot(data = mpg, aes(x = class)) +
      geom_bar()
    ```

-   21st finding outlier

    ```{r}
    library(dplyr)
    outlier <- data.frame(sex = c(1,2,1,3,2,1),     ##range 1~2
                           score = c(5,4,3,4,2,6))  ##range 1~5
    outlier

    table(outlier$sex)
    table(outlier$score)

    #if the vlaue of sex is 3, treat it as NA

    outlier$sex <- ifelse(outlier$sex == 3, NA, outlier$sex)
    outlier

    outlier$score <- ifelse(outlier$score >5, NA, outlier$score); outlier

    outlier %>% 
      filter(!is.na(sex) & !is.na(score)) %>% 
      group_by(sex) %>% 
      summarise(mean = mean(score))


    boxplot(mpg$hwy)
    boxplot(mpg$hwy)$stats

    mpg$hwy <- ifelse(mpg$hwy < 12 | mpg$hwy > 37, NA, mpg$hwy)
    table(is.na(mpg$hwy))

    mpg %>% 
      group_by(drv) %>% 
      summarise(mean_hwy = mean(hwy, na.rm = T)) %>% 
      arrange(desc(mean_hwy))

    ##practice

    mpg <- as.data.frame(ggplot2::mpg)
    #drv consists of 4, f, r

    names(mpg)
    mpg[c(10, 14, 58, 93), "drv"] <- "k"
    mpg[c(29, 43, 129, 203), "cty"] <- c(3, 4, 39, 42)

    #q1
    table(mpg$drv)
    mpg$drv <- ifelse(mpg$drv %in% c("4", "f", "r"), mpg$drv, NA)



    #q2
    boxplot(mpg$cty)
    boxplot(mpg$cty)$stats

    mpg$cty <- ifelse(mpg$cty < 9 | mpg$cty > 26, NA, mpg$cty)
      
    boxplot(mpg$cty)


    #q3
    mpg %>% 
      filter(!is.na(cty) & !is.na(drv)) %>% 
      group_by(drv) %>% 
      summarise(mean = mean(cty))
    ```

-   20th Review, process missing value

    ```{r}
    library(ggplot2)
    library(dplyr)
    data(midwest)
    mw <- midwest

    names(mw)
    mw <- mw %>% 
      mutate(rate_adol = (poptotal - popadults)/poptotal)

    mw %>% 
      mutate(rate_adol = (poptotal - popadults)/poptotal) %>% 
      select(rate_adol, county) %>% 
      arrange(desc(rate_adol)) %>% 
      head(5)
      

    mw <- mw %>% 
      mutate(grade = ifelse(rate_adol >= 0.4, "large", 
                            ifelse(rate_adol < 0.4 & rate_adol >= 0.3, "middle", "small")))


    table(mw$grade)


    mw %>% 
      mutate(rate_asia = popasian / poptotal) %>% 
      select(rate_asia, state, county) %>% 
      arrange(rate_asia)


    ### missing value
    df <- data.frame(sex = c("M", "F", NA, "M", "F"),
                     score = c(5, 4, 3, 4, NA))
    df

    is.na(df)
    table(is.na(df))
    table(is.na(df$sex))
    table(is.na(df$score))
    mean(df$scroe)
    sum(df$score)

    df %>% filter(is.na(score))
    df %>% filter(!is.na(score))

    df_nomiss <- df %>% filter(!is.na(score));df_nomiss
    sum(df_nomiss$score)

    df_nomiss <- df %>% filter(!is.na(score) & !is.na(sex))
    df_nomiss

    df_nomiss2 <- na.omit(df)
    df_nomiss2

    mean(df$score, na.rm = T)
    sum(df$score, na.rm = T)

    exam[c(3, 8, 15), "math"] <- NA
    exam

    exam %>% 
      summarise(mean_math = mean(math))

    exam %>% 
      summarise(mean_math = mean(math, na.rm = T))

    exam %>% summarise(mean_math = mean(math, na.rm = T),
                       sum_math = sum(math, na.rm = T), 
                       median_math = median(math, na.rm = T))

    mean(exam$math, na.rm = T)

    exam$math <- ifelse(is.na(exam$math), 55, exam$math)
    table(is.na(exam$math))

    exam


    mean(exam$math)

    mpg[c(65, 124, 131, 153, 212), "hwy"] <- NA

    table(is.na(mpg$hwy))
    table(is.na(mpg$drv))

    mpg %>% 
      filter(!is.na(hwy)) %>% 
      group_by(drv) %>% 
      summarise(mean = mean(hwy))

    mpg %>% 
      group_by(drv) %>% 
      summarise(mean = mean(hwy, na.rm = T))


    ```

-   19th Data pre-processing(summarise(), group_by()), left_join(), bind_rows())

    ```{r}
    library(dplyr)

    exam %>% 
      summarise(mean_math = mean(math))

    exam %>% 
      mutate(m_math = mean(math))

    exam %>% 
      group_by(class) %>% 
      summarise(mean_math = mean(math))

    exam %>% 
      group_by(class) %>% 
      summarise(mean_math = mean(math),
                sum_math = sum(math),
                median_math = median(math),
                n = n())

    mpg %>% group_by(manufacturer, drv) %>% 
      summarise(mean_cty = mean(cty)) %>% 
      head(10)

    library(ggplot2)
    data(mpg)

    mpg %>% 
      group_by(manufacturer) %>% 
      filter(class == "suv") %>% 
      mutate(tot = (cty+hwy)/2) %>% 
      summarise(mean_tot = mean(tot)) %>% 
      arrange(desc(mean_tot)) %>% 
      head(5)

    mpg %>% 
      group_by(class) %>% 
      summarise(m_cty = mean(cty)) %>% 
      arrange(desc(m_cty))

    mpg %>% 
      group_by(manufacturer) %>% 
      summarise(m_hwy = mean(hwy)) %>% 
      arrange(desc(m_hwy)) %>% 
      head(3)

    mpg %>% 
      filter(class == "compact") %>% 
      group_by(manufacturer) %>% 
      summarise(n = n())


    #merge => left_join
    test1 <- data.frame(id = c(1,2,3,4,5),
                        midterm = c(60, 70, 80, 90, 85)) 
    test2 <- data.frame(id = c(1,2,3,4,5),
                        final = c(70, 80, 65, 95, 80))
    test1;test2

    total <- left_join(test1, test2, by = "id"); total

    name <- data.frame(class = c(1,2,3,4,5),
                       teacher = c("kim", "lee", "park", "choi", "jung"))
    name
    exam

    exam_new <- left_join(exam, name, by = "class")
    exam_new


    #stacking data vertically
    group_a <- data.frame(id = c(1,2,3,4,5),
                          test = c(60, 70, 80, 90, 85))
    group_b <- data.frame(id = c(6, 7, 8, 9, 10),
                          test = c(70, 80, 65, 95, 80))
    group_a; group_b

    group_all <- bind_rows(group_a, group_b)
    group_all

    fuel <- data.frame(fl = c("c","d","e","p","r"),
                       price_fl = c(2.35, 2.38, 2.11, 2.76, 2.22))
    fuel

    mpg_fl <- left_join(mpg, fuel, by = "fl")
    head(mpg_fl,3)

    mpg_fl %>% 
      select(model, fl, price_fl) %>%
      group_by(model) %>% 
      head(10)
    ```

<!-- -->

-   18th Data pre-processing(arrange(), mutate())

```{r}

#ascending order

exam %>% 
  arrange(math)

#descending order

exam %>% 
  arrange(desc(math))

exam %>% 
  arrange(class, math)

mpg %>% 
  filter(manufacturer == "audi") %>% 
  arrange(desc(hwy)) %>% 
           head(5)

#mutate, derived variable
exam %>% 
  mutate(total = math + english + science) %>% 
  head

exam %>% 
  mutate(total = math + english + science,
         mean = (math + english + science)/3) %>% 
           head(3)

exam %>% 
  mutate(test = ifelse(science >= 60, "pass", "fail")) %>% 
  head

exam %>% 
  mutate(total = math + english + science,
         mean = (total/3),
         test = ifelse(mean > 90, "pass", "fail") )

exam %>% 
  mutate(total = math + english + science) %>% 
  arrange(desc(total)) %>% 
  head


mpg_total <- mpg %>% 
  mutate(total = cty + hwy,
         mean = total/2) %>% 
  arrange(desc(mean)) %>% 
  head(3)

mpg_total
```

-   17th Data pre-processing(filter(), select()

```{r}
library(dplyr)
library(haven)
library(readr)
exam <- read_csv("C:/Users/user/OneDrive - University of Illinois - Urbana/R/R/Doit_R-master/Doit_R-master/Data/csv_exam.csv")

head(exam, 6)

exam %>% filter(class == 1)

exam %>% filter(class == 2)

exam %>% filter(class != 1)
exam %>% filter(class != 3)

exam %>% filter(math > 50)
exam %>% filter(math < 50)
exam %>% filter(english >= 80)
exam %>% filter(english <= 80)

exam %>% filter(class == 1 & 
                  math >= 50)

exam %>% filter(class == 2, english >= 80)


#vertical bar, or 

exam %>% filter(math >= 90 | english >= 90)

exam %>%  filter(english < 90 | science < 50)

exam %>%  filter(class == 1 | class == 3 | class == 5)

#matching operator 
exam %>% filter(class %in% c(1,3,5))

class1 <- exam %>% 
  filter(class == 1)

class2 <- exam %>% 
  filter(class == 2)

mean(class1$math)
mean(class2$math)

library(ggplot2)
data(mpg)

names(mpg)

dspl4 <- mpg %>%
  filter(displ <= 4)

dspl5 <- mpg %>% 
  filter(displ >= 5)

head(dspl4$hwy, 6)
mean(dspl4$hwy)
mean(dspl5$hwy)


audi <- mpg %>% filter(
  manufacturer == "audi"
)

toyota <- mpg %>% 
  filter(manufacturer == "toyota")


mean(audi$cty)
mean(toyota$cty)

cfh <- mpg %>% 
  filter(manufacturer %in% c("chevolet", "ford", "hond"))
mean(cfh$hwy)



exam %>% select(math)
exam %>% select(english)
exam %>% select(class, math, english)

exam %>% select(-math)
exam %>% select(-math, -english)

exam %>% 
  filter(class == 1) %>% 
  select(english)


exam %>% 
  select(id, math) %>% 
  head

exam %>% 
  select(id, math) %>% 
  head(10)

data <- mpg %>% 
  select(class, cty)
  
head(data)

data_suv <- data %>% 
  filter(class == "suv"); mean(data_suv$cty) 
data_compact <- data %>% 
  filter(class == "compact"); mean(data_compact$cty) 
  

```

-   16th rename(), derived variable(conditional statements, nested conditional statements),

```{r}
df_raw <- data.frame(var1 = c(1,2,1),
                     var2 = c(2,3,2))
df_raw


library(dplyr)

df_new <- df_raw
df_new

df_new <- rename(df_new, v2 = var2)
df_new

df_new <- rename(df_new, v1 = var1)
df_new

library(ggplot2)
data(mpg)

A <- mpg
head(A)

A <- rename(A, city = cty, 
            highway = hwy)

names(A)
```

Derived variable

```{r}

df <- data.frame(var1 = c(4,3,8),
                 var2 = c(2, 6,1))
df

df$var_sum <- df$var1 + df$var2; df

df$var_mean <- df$var_sum/2
df

mpg$total <- (mpg$cty + mpg$hwy)/2
head(mpg)

mean(mpg$total)


```

conditional functions

```{r}
summary(mpg$total)
hist(mpg$total)
boxplot(mpg$total)

mpg$test <- ifelse(mpg$total > 20, "pass", "fail")

head(mpg, 20)
table(mpg$test)
qplot(mpg$test)

#Nested conditional statements 
# A >= 30, B 20~29, C < 20 

mpg$grade <- ifelse(mpg$total >= 30, "A",
                    ifelse(mpg$total <30 & mpg$total >=20, "B", "C"))

head(mpg$grade, 10)
table(mpg$grade)
qplot(mpg$grade)

mpg$grade2 <- ifelse(mpg$total >= 30, "A", 
                     ifelse(mpg$total >=25, "B",
                            ifelse(mpg$total >20, "C", "D")))
table(mpg$grade2)


```

## First Week (Sep. 9 \~ Sep. 13)

Basic R

Sep. 12

```{r}
library(readr)

exam <- read_csv("C:/Users/user/OneDrive - University of Illinois - Urbana/R/R/Doit_R-master/Doit_R-master/Data/csv_exam.csv") 

head(exam)  

head(exam, 10)

tail(exam)

tail(exam, 10)

View(exam)

dim(exam) #dimension

str(exam) #structure

summary(exam)

library(ggplot2)

data(mpg)
head(mpg)
tail(mpg)
View(mpg)
dim(mpg)
str(mpg)
summary(mpg)
```

Sep.9: Creating Dataframe

```{r}
english <- c(90, 80, 60, 70) #creat variable
english

math <- c(50, 60, 100, 20)
math
```

```{r}
df_midterm <- data.frame(english, math)

df_midterm
```

```{r}
class <- c(1,1,2,2)
class

df_midterm <- data.frame(english, math, class)
```

```{r}
mean(df_midterm$english)
mean(df_midterm$math)
```

```{r}
df_midterm <- data.frame(english = c(90, 80, 60, 70),
                         math = c(50, 60, 100, 20),
                         class = c(1,1,2,2))
df_midterm
```

```{r}
df_fruits <- data.frame(fruits = c("apple", "strawberry", "watermelon"),
                        price = c(1, 4, 7),
                        sales = c(24, 38, 13))
df_fruits

mean(df_fruits$price)
mean(df_fruits$sales)

mean(df_fruits$price * df_fruits$sales)
```
